Comparing vector-based and ACT-R memory models of tag retrieval:\protect\\User-customized hashtag/tag prediction on Twitter and StackOverflow

### 2013-11-07 writeups/proposal/proposal.tex
foo
Former-commit-id: 5918b46e49a1c152c0b77d4cb8325a1cdbaf4557 [formerly db83c52687909c3ec9764ec7d0b0b68a4774ac2f]
Former-commit-id: 02f679de7c067fc27e5a0db99ce52902406b8d76

\abstract{
  This research aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
  Two state-of-the-art cognitively-plausible memory models will be evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
  Experiments are designed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
  The user-customized tag-recommendation models developed with this research should be immediately useful for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
  Further, the planned performance modifications to the two memory models are aimed at the architectural level of each model.
  If these architectural modifications to two cognitively-plausible memory models significantly improve performance,
  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely highly correlated with one another.
It may be interesting to explore how model performance changes when one calculation is swapped out for another.
For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.

@@

The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on a large-scale real-world retrieval task,
[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.

>>>>

\abstract{
  This research aims to develop and compare tag-recommendation models for users choosing hashtags when composing a tweet on Twitter and choosing tags when asking a question on the StackOverflow site.
  Two state-of-the-art cognitively-plausible memory models will be evaluated on how accurate they predict a user's chosen tags: an ACT-R inspired Bayesian model and a random-permutation vector-based model.
  Experiments are designed to [1] discover the best way to leverage a user's prior hashtag use as a future predictor, [2] improve the models so that each model's strengths are incorporated into the other,
  and [3] formally evaluate and compare each model's performance when recommending hashtags on Twitter and tags on StackOverflow.
  The user-customized tag-recommendation models developed with this research should be immediately useful as foundations for tag-recommendation systems on the popular Twitter and StackOverflow social media sites.
  Further, the planned performance modifications to the two memory models are aimed at the architectural level of each model.
  If these architectural modifications to two cognitively-plausible memory models significantly improve performance,
  they may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

Vector-based and LSA retrieval models calculate chunk activation by correlating a representation of the current context with each memory item.
Bayesian retrieval models like ACT-R calculate activation by computing each chunk's posterior odd adjustment likelihood, given the context.
These two computations are certainly not equivalent, even though they are most likely correlated with one another.
It may be interesting to explore how model performance changes when one calculation is swapped out for another.
For example, a vector-based model could generate a compressed (reduced dimension) co-occurrence matrix, and then Bayesian calculations could be performed on that matrix to compute adjusted odd likelihoods.
Alternatively, a vector representation of the current context could be correlated with each column in a full co-occurrence matrix.
It is unlikely that the latter technique will improve performance, as \textcite{Landauer1997} has shown that performing correlations on uncompressed co-occurrence matrices does not perform well.
But it may be the case that using Bayesian computation on a compressed vector-based model improves performance.

@@

The end result of this research will be [1] a better understanding of how a user's past tagging history influences future tag use,
[2] a more thorough evaluation of the strengths, weaknesses, and differences between two cognitively-plausible memory models on large-scale real-world retrieval tasks,
[3] improvements to each memory model where the strengths of each model (word order and user prior likelihood) are incorporated into the other,
and [4] two efficient implementations of the models for making user-customized tag predictions for StackOverflow posts and Twitter tweets.
